\large{\bf Looking forwards and backwards through locally regulated populations }

### 2022-04-22 forwards-and-backwards.tex
.

%% %% %% %% %% %% %% %% %% %% %% %%
\subsection{Construction of individual lineages}

@@

Potential reproduction events occur at times $\tau_a + s$
for each point $(s, v, y, \kappa) \in \Pi_a$.
We say `potential' since if the resulting level is greater than $\lambda$,
the event does not happen.
If this is the $j^\text{th}$ point in $\Pi_a$,
the new label is $aj$ and the (potential) new level is
\begin{equation*}
    U_{aj}(\tau_a + s)
    =
    U_a(\tau_a + s)
    +
    \frac{
        v
    }{
        2 \lambda^{-1} \theta \gamma(X_a(\tau_a + s-), \eta(\tau_a + s-))
        r(X_a(\tau_a + s-) + y, \eta(\tau_a + s-))
    }.
\end{equation*}
If $U_{aj}(\tau_a+s) < \lambda$,
the new individual labeled $aj$ is produced,
and $\kappa$ determines which label is associated with the new location,
so
\begin{align*}
    X_{aj}(\tau_a + s)
    &=
    X_a(\tau_a+s-) + (1 - \kappa) y \\
    X_a(\tau_a+s)
    &=
    X_a(\tau_a+s-) + \kappa y .
\end{align*}
If $U_{aj}(\tau_a+s) \ge \lambda$, then $X_a$ is unchanged.
Recall that the parental \emph{individual} always retains their spatial location,
so that $\kappa = 0$ corresponds to the parent being assigned a new level,
and our line of descent switching to the offspring.

Consequently, 
the \emph{line of descent}
leading from the initial generation
to any label $a \in \labelspace$
can be explicitly constructed as follows.
Supposed that $a = (a_1, a_2, \ldots, a_n) \in \IN^n$,
and for $1 \le k \le n$ let $(a)_k = (a_1, a_2, \ldots, a_k)$.
Then, let $\zeta_a(t)$ denote the counting process that records
which label is present in the line of descent at time $t$,
i.e., $\zeta_a(t) = \max\{k \ge 1 : \tau_{(a)_k} \le t\}$.
This then lets us define $X_a(t)$ all the way back to $t=0$
by setting $X_a(t) = X_{(a)_{\zeta_a(t)}}(t)$ (and, $U_a(t)$ similarly).
Now, we can explicitly write down the evolution of $U_a(t)$ and $X_a(t)$.

First, for each $1 \le k < n$
let $(s_k, v_k, y_k, \kappa_k)$ be the $a_{k+1}^\text{th}$ point in $\Pi_{(a)_k}$
(this is the sequence of jumps taken by the line of descent).
Note that $s_k = \tau_{(a)_k}$, and also set $s_0 = 0$.
Let $\Gamma_a$ by the point process ``experienced by the line of descent leading to $a$'',
which is for $A \subset [0,\infty)^2 \times \IR^d \times \{0,1\}$
$$
\Gamma_a(A)
=
    \sum_{k=1}^n
    \Pi_{(a)_k}\left(A \cap [s_{k-1}, s_k) \times [0, \infty) \times \IR^d \times \{0,1\}\right) .
$$
Then for $0 \le t \le \sigma_a$,
\begin{align*}
    X_a(t)
    &=
    X_{a_1}(0)
    +
    \sum_{k=1}^{\zeta_a(t) - 1}
    y_k
    +
    \int_{
        [0, t) \times [0, \infty) \times \IR^d \times \{0, 1\}
    }
    \kappa y
    \Gamma_{a}(ds, dv, dy, d\kappa) .
\end{align*}

% Then for $0 \le t \le \sigma_a$,
% \begin{align*}
%     X_a(t)
%     &=
%     X_{a_1}(0)
%     +
%     \sum_{k=1}^{\zeta_a(t) - 1}
%     \left(
%     y_k
%     +
%     \int_{
%         [s_{k-1}, s_{k}) \times [0, \infty) \times \IR^d \times \{0, 1\}
%     }
%     \kappa y
%     \Pi_{(a)_k}(ds, dv, dy, d\kappa)
%     \right)
%     \\
%     &\qquad {}
%     +
%     \int_{
%         [s_{\zeta_a(t)}, t) \times [0, \infty) \times \IR^d \times \{0, 1\}
%     }
%     \kappa y
%     \Pi_{(a)_{\zeta_a(t)}}(ds, dv, dy, d\kappa) .
% \end{align*}

@@

%% %% %% %% %% %% %% %% %% %% %% %%
\subsection{Stochastic equations for movement of lines of descent}
    \label{sec:lines_of_descent}

>>>>

%% %% %% %% %% %% %% %% %% %% %% %%
\subsection{Construction tracking individual lines of descent}

\comment{This is a rewrite using Ulam-Harris labeling
of \ref{sec: Evolution of an Individual Particle}.}

@@

Potential reproduction events occur at times $\tau_a + s$
for each point $(s, v, y, \kappa) \in \Pi_a$.
We say `potential' since if the resulting level is greater than $\lambda$,
the event does not happen.
If this is the $j^\text{th}$ point in $\Pi_a$,
the new label is $aj$, the (potential) birth time is $\tau_{aj} = \tau_a + s$.
For convenience define
$$
    \ell(x, y, \eta, v)
    =
    \frac{
        v
    }{
        2 \lambda^{-1} \theta \gamma(x, \eta) r(x + y, \eta)
    } ,
$$
and using this, the (potential) new level is
\begin{equation*}
    U_{aj}(\tau_{aj})
    =
    U_a(\tau_{aj})
    +
    \ell(X_a(\tau_{aj}-), y, \eta(\tau_{aj}-), v) .
\end{equation*}
If $U_{aj}(\tau_{aj}) < \lambda$,
the new individual labeled $aj$ is produced,
and $\kappa$ determines which label is associated with the new location,
so
\begin{align*}
    X_{aj}(\tau_a + s)
    &=
    X_a(\tau_a+s-) + (1 - \kappa) y .
\end{align*}
On the other hand
if $U_{aj}(\tau_a+s) \ge \lambda$, then $X_a$ is unchanged and $X_{aj}$ is undefined,
so
\begin{align*}
    X_a(\tau_a+s)
    &=
    X_a(\tau_a+s-) + \kappa y \mathbbm{1}_{U_{aj}(\tau_{aj}) < \lambda} .
\end{align*}
Recall that the parental \emph{individual} always retains their spatial location,
so that $\kappa = 0$ corresponds to the parent being assigned a new level,
and our line of descent switching to the offspring.

Consequently, 
the \emph{line of descent}
leading from the initial generation
to any label $a \in \labelspace$
with $U_a(\tau_a) < \lambda$
can be explicitly constructed as follows.
Supposed that $a = (a_1, a_2, \ldots, a_n) \in \IN^n$,
and for $1 \le k \le n$ let $(a)_k = (a_1, a_2, \ldots, a_k)$.
Then, let $\zeta_a(t)$ denote the counting process that records
which label is present in the line of descent at time $t$,
i.e., $\zeta_a(t) = \max\{k \ge 1 : \tau_{(a)_k} \le t\}$.
This then lets us define $X_a(t)$ all the way back to $t=0$
by setting $X_a(t) = X_{(a)_{\zeta_a(t)}}(t)$ (and, $U_a(t)$ similarly).
Now, we can explicitly write down the evolution of $U_a(t)$ and $X_a(t)$.

\comment{Do we want subscript $a$s here, on e.g., $\zeta$?
I think not, because this is all given a particular $a$?}

First, for each $1 \le k < n$
let $(s_k, v_k, y_k, \kappa_k)$ be the $a_{k+1}^\text{th}$ point in $\Pi_{(a)_k}$
(this is the sequence of jumps taken by the line of descent).
Note that $s_k = \tau_{(a)_k}$, and also set $s_0 = 0$.
Let $\Gamma_a$ by the point process ``experienced by the line of descent leading to $a$'',
which is easier to define if we treat point processes as sets for a moment,
defining
\begin{align*}
    \Gamma_a
    &=
    \bigcup_{k=1}^n 
    \bigg\{
        (s, v, y, \kappa) \in \Pi_{(a)_k} \; : \;
        \tau_{(a)_{k-1}} \le s < \tau_{(a)_{k-1}} 
    \\& \qquad \qquad 
        \text{ and }
        U_{(a)_k}(\tau_{(a)_{k-1}} + s) + \ell( X_{(a)_k}(\tau_{(a)_{k-1}} + s-), y, \eta(\tau_{(a)_{k-1}} + s-), v) < \lambda
    \bigg\}
\end{align*}
\comment{TODO include the switch points}

which is for $A \subset [0,\infty)^2 \times \IR^d \times \{0,1\}$
\comment{This is wrong, as it includes non-viable individuals.}
\begin{align*}
\Gamma_a(A)
    &=
    \sum_{k=1}^n
    \Pi_{(a)_k}\left(A \cap [s_{k-1}, s_k) \times [0, \infty) \times \IR^d \times \{0,1\}\right) 
    \\ &\qquad {}
    + \sum_{k=1}^{n-1}
      \delta_{(s_k, v_k, y_k, 1 - \kappa_k)}(A) .
\end{align*}
Note that we have included the points at which the line of descent's level jumps,
flipping $\kappa$ at these points,
and as a result $\Gamma_a$ has the same distribution as $\Pi_a$
restricted to $[0, \sigma_a)$.
Then for $0 \le t \le \sigma_a$,
\begin{align} \label{eqn:X_line_of_descent}
    X_a(t)
    &=
    X_{a_1}(0)
    +
    \int_{
        [0, t) \times [0, \infty) \times \IR^d \times \{0, 1\}
    }
    \kappa y
    \mathbbm{1}_{
        U_{(a)_{\zeta(s)}}(s) + \ell( X_{(a)_{\zeta(s)}}(s-), y, \eta(s-), v) < \lambda
    }
    \Gamma_{a}(ds, dv, dy, d\kappa) 
\end{align}
and
\begin{align} \label{eqn:U_line_of_descent}
    U_a(t)
    &=
    U_{a_1}(0)
    +
    \sum_{k=1}^{\zeta_a(t) - 1}
    \frac{
        v_k
    }{
        2 \lambda^{-1} \theta \gamma(X_{(a)_k}(s_k-), \eta(s_k-))
        r(X_{(a)_k}(s_k-) + y_k, \eta(s_k-))
    }
    \\ & \qquad {}
    +
    \int_0^t
    \bigg(
        \frac{\theta}{\lambda} \gamma(X_a(s),\eta(s))
        \int_{\IR^d} r(z,\eta(s)) q(X_a(s),dz) U_a(s)^2
    \\ & \qquad\qquad  \qquad {}
        -
        b(X_a(s),\eta(s)) U_a(s)
    \bigg)
    dt .
\end{align}

%% %% %% %% %% %% %% %% %% %% %% %%
\subsection{Limiting processes for lines of descent}

\comment{This is a rewrite of \ref{sec:lines_of_descent} following on the previous section}

The previous section constructs the lookdown process
in a way that makes individual lines of descent converge pointwise as $\lambda \to \infty$,
using the same underlying Poisson processes $\{\Pi_a\}_{a \in \labelspace}$.
To see this, first note that if the Poisson processes are fixed
then the set of events to which a given label $a \in \labelspace$ is associated
is also fixed -- this is the sequence $(s_k, v_k, y_k, \kappa_k)$ associated with a given label $a$.
Therefore, the processes $\Gamma_a$ are also unchanged.
First, we clearly need that the spatial measures $\eta$ converge.
\comment{(hm, no notation for $\eta$ depending on $\lambda$ here; need to introduce this here?)}
Now, consider equations~\eqref{eqn:X_line_of_descent} and~\eqref{eqn:U_line_of_descent}.

Equation~\eqref{eqn:X_line_of_descent} simply says that $X_a$ follows a random walk

@@

%% %% %% %% %% %% %% %% %% %% %% %%
\subsection{PREVIOUS VERSION OF Stochastic equations for movement of lines of descent}
    \label{sec:lines_of_descent}
